Measuring Directionality in Double-Beta Decay and Neutrino Interactions with Kiloton-Scale Scintillation Detectors

### 2013-07-22 Direction_v6_submit.tex
The files submitted to the arXiv. --Lindley

\title{Measuring Directionality in Double Beta Decay and Neutrino Interactions with Kiloton-Scale Scintillation Detectors}

@@

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.4]{graphs/geometry_plot_labels.pdf}
        \caption[]{The detector geometry and coordinate system.
        The radial rays (green lines) are photons emitted by two back-to-back electrons with 1.4~MeV each
        (equally divided energy of $^{116}$Cd~$0\nu\beta\beta$ decay). The electrons originate at
        the center of the sphere with initial directions along the x
        and -x-axis. Only Cherenkov photons are drawn to illustrate the
        directionality of the event. \label{detector_view}}
        \end{center}
\end{figure}

@@

\begin{figure*}[tbh]
\begin{center}
        \subfigure[ ~Default
        simulation.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
        \subfigure[ ~Increased TTS
        (1.28~ns).]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
        \subfigure[ ~Red-sensitive
        photocathode.]{\includegraphics[scale=0.295]{graphs/6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
        \caption[]{Photo-electron (PE) arrival times after application
        of the transit-time spread (TTS) for the simulation
        of 1000 electrons (5~MeV) with different values of the TTS and
        wavelength response. PEs from
        Cherenkov light (black, solid line) and scintillation light
        (red, dotted line) are
        compared. The dash-dotted vertical line illustrates a time cut at
        34.0~ns. (a) Default simulation: bialkali photocathode and TTS =
        0.1~ns ($\sigma$). After the 34.0~ns time cut we get 171~PEs
        from scintillation and 108~PEs from Cherenkov light. (b)
        Default simulation settings except for TTS = 1.28~ns (KamLAND
        17 in. PMTs). After the 34.0~ns time cut we get 349~PEs from
        scintillation and 88~PEs from Cherenkov light. (c) Default
        simulation settings except for a GaAsP photocathode. After the
        34.0~ns time cut we get 226~PEs from scintillation and 229~PEs
        from Cherenkov light. \label{time_plots_comparison}}
        \end{center}
\end{figure*}

@@

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.40]{graphs/cos_psi_34_h.pdf}
        \caption[]{The angular distribution of photoelectron hits
        relative to the original electron direction, $\cos(\Psi) =
        x_{hit}/\vec{r}_{hit}$. The sample consists of 1000 events
        with a 5~MeV electron produced at the detector center. Default
        simulation settings are used and both Cherenkov and
        scintillation light are included.} 
        \label{Cherenkov_cone}
        \end{center}
\end{figure}

@@

\begin{figure*}[tbh]
        \begin{center}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirX_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirY_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hDirZ_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxX_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxY_v1.pdf}}
        %%\subfigure{\includegraphics[scale=0.295]{graphs/hVtxZ_v1.pdf}}
        \subfigure{\includegraphics[scale=0.4]{graphs/hDir_34ns.pdf}}
        \subfigure{\includegraphics[scale=0.4]{graphs/hVtx_34ns.pdf}}
        \caption[]{\label{fig:reco} (Left) The reconstructed direction,
        $(\frac{p_x}{|\vec{p}|}, \frac{p_y}{|\vec{p}|},
        \frac{p_z}{|\vec{p}|})$, for the simulation of 1000 electrons
        (5~MeV). In the simulation the electrons are produced along the
        x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0), and originate
        from the center of the 6.5m-radius detector, $\vec{r}$ =
        (0,0,0). Only photons with arrival time of $t<$ 34.0~ns are used
        in the reconstruction. The quantum efficiency of the bialkali
        photocathode is taken into account. (Right) The reconstructed
        vertex position, $(x,y,z)$, for the same simulation.}

@@

\section{Acknowledgments}
The authors thank Andrew Blake at University of Cambridge for his work
authoring the WCSimAnalysis code. The authors thank the neutrino
reconstruction group at Iowa State, particularly Mayly Sanchez, Ioana
Anghel, and Tian Xin, for their continued work in developing the
WCSimAnalaysis algorithms and for their insights and expertise
regarding issues related to Cherenkov reconstruction with
fast-timing. The authors also thank Micheal Smy for his development of
the quadruplet-based vertex-finding method. L. Winslow would like to
thank J.M. Conrad for many useful discussion on the topic, and
K. Arisaka for discussions on the possible reach of traditional PMTs
and the characteristics of HPDs. C. Aberle and L. Winslow are
supported by funds from University of California Los Angeles. The work
at the University of Chicago is partially supported by DOE
contract DE-SC0008172 and NSF grant PHY-1066014. Matthew Wetstein gratefully
acknowledges support by the Grainger Foundation.

>>>>

\title{Measuring Directionality in Double-Beta Decay and Neutrino Interactions with Kiloton-Scale Scintillation Detectors}

@@

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.4]{geometry_plot_labels.pdf}
        \caption[]{The detector geometry and coordinate system.
        The radial rays (green lines) are photons emitted by two back-to-back electrons with 1.4~MeV each
        (equally divided energy of $^{116}$Cd~$0\nu\beta\beta$ decay). The electrons originate at
        the center of the sphere with initial directions along the x
        and -x-axis. Only Cherenkov photons are drawn to illustrate the
        directionality of the event. \label{detector_view}}
        \end{center}
\end{figure}

@@

\begin{figure*}[tbh]
\begin{center}
        \subfigure[ ~Default
        simulation.]{\includegraphics[scale=0.295]{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_TIME.pdf}}
        \subfigure[ ~Increased TTS
        (1.28~ns).]{\includegraphics[scale=0.295]{6p5Meter_5MeVElectrons_Bialkali_KamlandScintSpec_1p28nsTTS_TIME.pdf}}
        \subfigure[ ~Red-sensitive
        photocathode.]{\includegraphics[scale=0.295]{6p5Meter_5MeVElectrons_RedSensitiveQE_KamlandScintSpec_TIME.pdf}}
        \caption[]{Photo-electron (PE) arrival times after application
        of the transit-time spread (TTS) for the simulation
        of 1000 electrons (5~MeV) with different values of the TTS and
        wavelength response. PEs from
        Cherenkov light (black, solid line) and scintillation light
        (red, dotted line) are
        compared. The dash-dotted vertical line illustrates a time cut at
        34.0~ns. (a) Default simulation: bialkali photocathode and TTS =
        0.1~ns ($\sigma$). After the 34.0~ns time cut we get 171~PEs
        from scintillation and 108~PEs from Cherenkov light. (b)
        Default simulation settings except for TTS = 1.28~ns (KamLAND
        17 in. PMTs). After the 34.0~ns time cut we get 349~PEs from
        scintillation and 88~PEs from Cherenkov light. (c) Default
        simulation settings except for a GaAsP photocathode. After the
        34.0~ns time cut we get 226~PEs from scintillation and 229~PEs
        from Cherenkov light. \label{time_plots_comparison}}
        \end{center}
\end{figure*}

@@

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.40]{cos_psi_34_h.pdf}
        \caption[]{The angular distribution of photoelectron hits
        relative to the original electron direction, $\cos(\Psi) =
        x_{hit}/\vec{r}_{hit}$. The sample consists of 1000 events
        with a 5~MeV electron produced at the detector center. Default
        simulation settings are used and both Cherenkov and
        scintillation light are included.} 
        \label{Cherenkov_cone}
        \end{center}
\end{figure}

@@

\begin{figure*}[tbh]
        \begin{center}
        \subfigure{\includegraphics[scale=0.4]{hDir_34ns.pdf}}
        \subfigure{\includegraphics[scale=0.4]{hVtx_34ns.pdf}}
        \caption[]{\label{fig:reco} (Left) The reconstructed direction,
        $(\frac{p_x}{|\vec{p}|}, \frac{p_y}{|\vec{p}|},
        \frac{p_z}{|\vec{p}|})$, for the simulation of 1000 electrons
        (5~MeV). In the simulation the electrons are produced along the
        x-axis, $\frac{\vec{p}}{|\vec{p}|}$ = (1,0,0), and originate
        from the center of the 6.5m-radius detector, $\vec{r}$ =
        (0,0,0). Only photons with arrival time of $t<$ 34.0~ns are used
        in the reconstruction. The quantum efficiency of the bialkali
        photocathode is taken into account. (Right) The reconstructed
        vertex position, $(x,y,z)$, for the same simulation.}

@@

\section{Acknowledgments}
The authors thank Andrew Blake at University of Cambridge for his work
authoring the WCSimAnalysis code. The authors thank the neutrino
reconstruction group at Iowa State, particularly Mayly Sanchez, Ioana
Anghel, and Tian Xin, for their continued work in developing the
WCSimAnalaysis algorithms and for their insights and expertise
regarding issues related to Cherenkov reconstruction with
fast-timing. The authors also thank Micheal Smy for his development of
the quadruplet-based vertex-finding method. L. Winslow would like to
thank Janet Conrad for many useful discussion on the topic, and
Katsushi Arisaka for discussions on the possible reach of traditional PMTs
and the characteristics of HPDs. C. Aberle and L. Winslow are
supported by funds from University of California Los Angeles. The work
at the University of Chicago is partially supported by DOE
contract DE-SC0008172 and NSF grant PHY-1066014. Matthew Wetstein gratefully
acknowledges support by the Grainger Foundation.
